Technical Dimensions of Programming Systems

### 2022-06-01 prog-2022/prog22-master.tex
Submission-ready

Technical dimensions capture characteristics that may be studied, compared and advanced independently. This makes it possible to talk about programming systems in a way that can be shared and constructively debated rather than relying solely on personal impressions. We outline two concrete ways of using our framework. First, we show how to use it to analyze a recently developed novel programming system, and, second, we use it to identify an interesting unexplored point in the design space of programming systems. By providing foundations for more systematic research, we can help programming systems researchers to stand, at last, on the shoulders of giants.

>>>>

Technical dimensions capture characteristics that may be studied, compared and advanced independently. This makes it possible to talk about programming systems in a way that can be shared and constructively debated rather than relying solely on personal impressions. We outline two concrete ways of using our framework. First, we show how it can analyze a recently developed novel programming system. Then, we use it to identify an interesting unexplored point in the design space of programming systems. By providing foundations for more systematic research, we can help programming systems researchers to stand, at last, on the shoulders of giants.
